---
title: "Homework 1"
author: "Matthew Hoctor, PharmD"
date: last-modified

quarto::html_document:
  theme: cerulean
  highlight: github
  
toc: true
toc-depth: 4
toc-location: left
toc-title: Contents
  
code-fold: show
code-overflow: wrap
code-tools: true
code-link: true

editor: source
---

```{r setup, include = FALSE}
# load libraries:
library(tidyverse)
library(rethinking)
# library(data.table)
```

# 0: EASY problems

The EASY problems at the end of Chapters 1, 2 and 3 (2nd edition).

## CH2

### 2E1

Which of the expressions below correspond to the statement: the probability of rain on Monday? 
(1) Pr(rain) 
(2) Pr(rain|Monday) 
(3) Pr(Monday|rain) 
(4) Pr(rain, Monday)/ Pr(Monday) 

$Pr(rain|Monday) = \frac{Pr(rain, Monday)}{Pr(Monday)}$, therefore: 2 & 4

### 2E2

Which of the following statements corresponds to the expression: Pr(Monday|rain)? 
(1) The probability of rain on Monday. 
(2) The probability of rain, given that it is Monday. 
(3) The probability that it is Monday, given that it is raining. 
(4) The probability that it is Monday and that it is raining. 

3

### 2E3

Which of the expressions below correspond to the statement: the probability that it is Monday, given that it is raining? 
(1) Pr(Mondayjrain) 
(2) Pr(rainjMonday) 
(3) Pr(rainjMonday) Pr(Monday) 
(4) Pr(rainjMonday) Pr(Monday)/ Pr(rain) 
(5) Pr(Mondayjrain) Pr(rain)/ Pr(Monday)

$\frac{Pr(rain|Monday)*Pr(Monday) }{Pr(rain)} = \frac{Pr(rain, Monday)}{Pr(rain)} = Pr(Monday|rain)$ therefore 1 & 4

### 2E4

The Bayesian statistician Bruno de Finetti (1906–1985) began his book on probability theory with the declaration: “PROBABILITY DOES NOT EXIST.” The capitals appeared in the original, so I imagine de Finetti wanted us to shout this statement. What he meant is that probability is a device for describing uncertainty from the perspective of an observer with limited knowledge; it has no objective reality. Discuss the globe tossing example from the chapter, in light of this statement. What does it mean to say “the probability of water is 0.7”?

In the globe tossing example the observer has devised a procedure to estimate the proportion of the surface of a globe which is covered in water by tossing it repeatedly and observing and recording the surface under their right index finger.  In light of Finetti's view of probability, the statement "the probability of water is 0.7" means that the observer has observed that their right index finger landed on water in their sample of tosses with a probability of 0.7.  The abstraction of the concept of water is descriptive of the substance the observer is trying to measure; similarly the abstraction of probability is descriptive of the proportion of the globe covered in water, given the assumptions of the estimation procedure, and limited information from the oberver's sample.  Thus when Finetti declares that “PROBABILITY DOES NOT EXIST” the declaration serves to underscore that probability is a descriptive abstraction, and not a thing in and of itself.

## CH3

These problems use the samples from the posterior distribution for the globe tossing example.  This code will give you a specific set of samples, so that you can check your answers exactly.

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```


### 3E1

How much posterior probability lies below p = 0.2?

### 3E2

How much posterior probability lies above p = 0.8?

### 3E3

How much posterior probability lies between p = 0.2 and p = 0.8?

### 3E4

20% of the posterior probability lies below which value of p?

### 3E5

20% of the posterior probability lies above which value of p?

### 3E6

Which values of p contain the narrowest interval equal to 66% of the posterior probability?

### 3E7

Which values of p contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval?



# 1

Suppose the globe tossing data (Lecture 2, Chapter 2) had turned out to be 3 water and 11 land. Construct the posterior distribution.

```{r}
curve( dbeta(x,3+1,11+1) , from=0 , to=1 , xlab="p" )
```


# 2

Using the posterior distribution from 1, compute the posterior predictive
distribution for the next 5 tosses of the same globe. I recommend you use
the sampling method.

# 3 OPTIONAL

OPTIONAL. This problem is an optional challenge for people who are
taking the course for a second or third time. Suppose you observe W = 7
water points, but you forgot to write down how many times the globe was
tossed, so you don’t know the number of land points L. Assume that p = 0.7
and compute the posterior distribution of the number of tosses N. Hint: Use
the binomial distribution.

# Session Info

```{r}
sessionInfo()
```